---
title: "Basket Analysis for Rob's Market"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, nobreak=TRUE)
library(lemon)
library(plyr)
library(cleaner)
data <- read.csv('MarketBasket.csv')
knit_print.data.frame <- lemon_print
```

This report contains data collected on 1000 customer purchase items on: Bread, Jelly, Peanut Butter. It will examine some probability results and frequency breakdowns of items.



```{r caption='Brief preview of Data',render=lemon_print}
head(data)
```

## Frequency breakdown by items and their type
```{r Freq, caption='Frequency of Bread', render=lemon_print}
freq(x=clean_factor(data$Bread), 
     markdown=TRUE,
     title='Frequencies of Bread Type Sold',
     header=FALSE,
     quote=FALSE)
```

```{r, caption='Frequency of Jelly Type', render=lemon_print}
freq(x=clean_factor(data$Jelly), 
     markdown=TRUE,
     title='Frequencies of Jelly Type Sold',
     header=FALSE,
     quote=FALSE)
```

```{r, caption='Frequency of Peanut Butter Type', render=lemon_print}
freq(x=clean_factor(data$Peanut.Butter), 
     markdown=TRUE,
     title='Frequencies of Peanut Butter Type Sold',
     header=FALSE,
     quote=FALSE)
```

## Probability Analysis

1. The probability that a random customer does not purchase any of the three
products (bread, jelly, or peanut butter).
```{r 1}
none <- nrow(data[data$Bread == 'none' & data$Jelly == 'none' & data$Peanut.Butter == 'none',])
noneProb <- none/nrow(data)
noneProb
```

2. The probability that a random customer purchases white bread.
```{r 2, echo=FALSE}
white_bread_probability <- sum(data$Bread == 'white')/nrow(data)
white_bread_probability
```

3. The probability that a random customer purchases wheat bread.
```{r 3}
wheat_prob <- sum(data$Bread == 'wheat')/nrow(data)
wheat_prob
```

4. The probability that a random customer purchases grape jelly given that he or she purchases white bread.

```{r 4}
jelly_given_white <- sum(data$Jelly == 'grape' & data$Bread == 'white')/sum(data$Bread == 'white')
jelly_given_white
```
5. The probability that a random customer purchases strawberry jelly given that he or she purchases white bread.

```{r 5}
straw_given_white <- sum(data$Jelly == 'strawberry' & data$Bread == 'white')/sum(data$Bread == 'white')
straw_given_white
```
6. The probability that a random customer purchases creamy peanut butter given that he or she purchases white bread.
```{r 6}
cream_given_white <- sum(data$Peanut.Butter == 'creamy' & data$Bread == 'white')/sum(data$Bread == 'white')
cream_given_white
```

7. The probability that a random customer purchases natural peanut butter given that he or she purchases white bread.
```{r 7}
nat_given_white <- sum(data$Peanut.Butter == 'natural' & data$Bread == 'white')/sum(data$Bread == 'white')
nat_given_white
```

8. The probability that a random customer purchases creamy peanut butter given that
he or she purchases wheat bread.
```{r 8}
cream_given_wheat <- sum(data$Peanut.Butter == 'creamy' & data$Bread == 'wheat')/sum(data$Bread == 'wheat')
cream_given_wheat
```
9. The probability that a random customer purchases natural peanut butter given that he or she purchases wheat bread.
```{r 9}
nat_given_wheat <- sum(data$Peanut.Butter == 'natural' & data$Bread == 'wheat')/sum(data$Bread == 'wheat')
```

10. The probability that a random customer purchases white bread, grape jelly, and
creamy peanut butter.
```{r 10}
all <- nrow(data[data$Bread == 'white' & data$Jelly == 'grape' & data$Peanut.Butter == 'creamy',])
allprob <- all/nrow(data)
allprob
```

## Analysis

Looking through the probability analysis we see grape jelly has a high probability of being sold when white bread is purchased. It would be a good idea to display these two products to help drive sales. 

